---
title: "Using the Mortgage Estimator"
output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Intoduction

When considering buying a house or condominium, it is important to understand how your monthly payment will be affected by the terms of your mortgage and other items, such as property taxes and insurance.  This tool lets you explore how changes these parameters affects your total monthly housing payment.

Most mortagages in the United States are self-amortizing.  If the interest rate on the the mortgage is fixed, then the payments will be constant over the life of the loan.  However, the portions of the loan paying principal and interest change over time.  At first, nearly all of the payment pays interest, and very little of the principal.  Near the end of the loan, nearly all of the payment pays principal, and very little interest.

## Formula

This formula describes how the monthly payment is calculated from the loan's parameters.

$$Payment = P * (\frac{J}{(1-(1+J)^{-N}})$$

where<br/> 
P = the principal of the loan,<br/>
J = the interest rate divided by 12, and <br/>
N = the length of the mortgage in months.

To this is added the monthly property tax and property insurance amounts.

## Instructions

To use this tool, simply move the sliders to match the paramters of your mortgage.  If, for example, your<br/>

principal = $400,000,<br/>
down payment = 20%, <br/>
interest rate = 3.25%,<br/>
term of mortgage = 30 years,<br/>
annual property tax = $6000, and <br/>
annual property insurance = $3000, <br/><br/>

the combined monthly payment would be
```{r}
400000*(1-.2)*((3.25/12/100)/(1-(1+(3.25/12/100))^-360)) + 6000/12 + 3000/12
```

## 

Enjoy!<br/>

<a href>"https://thowing.shinyapps.io/mortgage_tool/"
